---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

# contrastable

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/tsostarics/contrastable/branch/main/graph/badge.svg?token=PW2NOWO8NE)](https://codecov.io/gh/tsostarics/contrastable?branch=main)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check](https://github.com/tsostarics/contrastable/workflows/R-CMD-check/badge.svg)](https://github.com/tsostarics/contrastable/actions)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.11869428.svg)](https://doi.org/10.5281/zenodo.11869428)

<!-- badges: end -->

This package provides utilities to set different common contrast coding schemes
for use with regression models.


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tsostarics/contrastable", build_vignettes = TRUE)
```

## Citation

To cite contrastable in publications, please use

Sostarics, T. (2024). contrastable: Contrast Coding Utilities in R. R package version 0.2.0.

A BibTeX entry for LaTeX users is

```
@Manual{,
    author = {Thomas Sostarics},
    title = {{contrastable}: Contrast Coding Utilities in {R}},
    year = {2024},
    note = {R package version 0.2.0},
    url = {https://github.com/tsostarics/contrastable},
    doi = {10.5281/zenodo.11869427},
  }
```

In-text citations should reference the package and ideally which contrast 
functions were used to avoid ambiguity.

BAD: "We sum code our variables (Sostarics 2024)"

GOOD: "We use the `sum_code()` function from the contrastable package (Sostarics 2024) for our variables."

Other examples:

 - Contrasts were set using the `scaled_sum_code()` function from the contrastable
 package (Sostarics 2024).
 - We use the contrastable package (Sostarics 2024) for contrast coding our
 categorical variables; details of the contrasts for each model are provided in
 the appendix. (be sure to do the latter!)
 - Below are the contrast matrices returned by the `helmert_code()` and
 `scaled_sum_code()` functions from the contrastable R package (Sostarics 2024)
 - We use the `set_contrasts()` and `sum_code()` functions from the contrastable package
 (Sostarics 2024) to sum code (+1/-1) our variables
 
I also recommend writing out, potentially in a footnote, what the comparisons are.

GOOD: "We use the contrastable package's `sum_code()` function (+1/-1, Sostarics 2024)."

BETTER: "We use the contrastable package's `sum_code()` function (+1/-1, Sostarics 2024).
This contrast scheme encodes differences between each comparison level and the grand mean."


# Example

Here is a simple example showing how to set particular factors to a specific
contrast scheme. 

```{r}
library(contrastable)
my_data <- mtcars
my_data$gear <- ordered(my_data$gear) # Set as ordered factor in dataframe
```

We can use `glimpse_contrasts` to get information about
the factors and diagnostics about the scheme we have set.

```{r, message = TRUE}
# Specify the contrast schemes we want, factor conversion done automatically
# Set reference level with + and intercept with *
contrast_schemes <- list(
  cyl ~ scaled_sum_code + 6,
  carb ~ helmert_code,
  vs ~ treatment_code + 1
)

# Get information about our contrasts, even those we didn't explicitly set
# (gear is ordered, and so uses contr.poly by default)
glimpse_contrasts(my_data,
  contrast_schemes,
  add_namespace = TRUE,
  all.factors = TRUE
) |>
  knitr::kable()
```


`enlist_contrasts` can be used to generate a named list of contrasts
that can be used in the `contrasts` argument of various modeling functions.

```{r, message = TRUE}
# Get a list of the contrasts we've explicitly set
enlist_contrasts(mtcars, contrast_schemes)
```

`set_contrasts` can be used to set the contrasts onto the dataframe itself,
which is needed when a modeling function lacks a `contrasts` argument.

```{r, message = TRUE}
# Set contrasts to dataframe itself
my_data <- set_contrasts(my_data, contrast_schemes)
MASS::fractions(contrasts(my_data$carb))
```

Functions will give various messages and warnings such as:

 - When non-factor columns are coerced to factors
 - When ordered factors are set to non-polynomial contrasts
 - When a factor with only one level is detected
 - When a factor is not explicitly set
 - and more


Detailed usage is available in the `contrasts` vignette with 
`vignette('contrasts', 'contrastable')`.
